---
title: "THE AMOUNT OF DETAILS IN TRUTHS AND LIES ABOUT PAST AND FUTURE WEEKEND ACTIVITIES"
author: "Sally A.M. Hogenboom - 11377909"
date: "16-12-2017"
output:
  pdf_document: default
  html_document: default
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# set WD
```


# Available Data

+ File 1 - [LieDetection_Data.csv](https://github.com/SHogenboom/OnlineAppendix--LieDetection2017/blob/master/LieDetection_Data.csv)
    + Contains demographics, veracity judgement, exclusion criteria, transcriptions,
  confidence of deceit, frequency of lie.
+ File 2 - [LieDetection_NER_processed.csv](https://github.com/SHogenboom/OnlineAppendix--LieDetection2017/blob/master/LieDetection_NER_processed.csv)
    + Transcriptions in long format -> allows for Named Entity Recognition 
  (NER;https://spacy.io)
    + Unique and Non-Unique Frequency counts of the NERs...
          + Person
          + Location
          + Time
          + Facility
          + GPE
          + Product
          + Date
          + Money

```{R Reading in off data, echo = FALSE}
# Read in file 1 - demographics
  file1 <- read.csv(file = "LieDetection_Data.csv", header = TRUE, sep = ";")

# Read in file 2 - processed statements
  file2 <- read.csv(file = "LieDetection_NER_processed.csv", header = TRUE, sep = ";")
  file2 <- file2[,-1] # remove empty first column

# Combine files - NO EXCLUSIONS
  # include only data from non-unique entities
  # if wish to include unique entitie counts -> include columns [13:21]
  Data_All <- cbind(file1, file2[1:77, 4:12], file2[78:154, 4:12])
  
# Rename columns from file 2 to have unique values
  # Saturday 
    colnames(Data_All)[21:29] <- c("Sat_Veracity", "Sat_Person_NU", 
                           "Sat_Location_NU", "Sat_Time_NU", 
                           "Sat_Facility_NU", "Sat_GPE_NU", 
                           "Sat_Product_NU", "Sat_Date_NU", "Sat_Money_NU")
  
  # Sunday 
    colnames(Data_All)[30:38] <- c("Sun_Veracity", "Sun_Person_NU", 
                           "Sun_Location_NU", "Sun_Time_NU", 
                           "Sun_Facility_NU", "Sun_GPE_NU", 
                           "Sun_Product_NU", "Sun_Date_NU", "Sun_Money_NU")
  # Others
    colnames(Data_All)[11] <- "OrderOfLie"
    colnames(Data_All)[2] <- "Time"

  # Calculate total amount of NER in statements
  # Sum the frequency counts of the specific NERs into a total value of Non-Unique 
    # ... entities.
      Data_All$Total_Sat <- rowSums(Data_All[,c("Sat_Person_NU", "Sat_Location_NU", 
                                      "Sat_Time_NU", "Sat_Facility_NU", "Sat_GPE_NU", 
                                      "Sat_Product_NU", "Sat_Date_NU", "Sat_Money_NU")])
    
      Data_All$Total_Sun <- rowSums(Data_All[, c("Sun_Person_NU", "Sun_Location_NU", 
                                       "Sun_Time_NU", "Sun_Facility_NU", 
                                       "Sun_GPE_NU", "Sun_Product_NU", "Sun_Date_NU", 
                                       "Sun_Money_NU")])
      
# Calculate Total NERs Divided for Truths and Lies
  # create new columns to store output
    Data_All$Person_Truth <- NaN
    Data_All$Time_Truth <- NaN
    Data_All$GPE_Truth <- NaN
    Data_All$Date_Truth <- NaN
    Data_All$Total_Truth <- NaN
  
    Data_All$Person_Lie <- NaN
    Data_All$Time_Lie <- NaN
    Data_All$GPE_Lie <- NaN
    Data_All$Date_Lie <- NaN
    Data_All$Total_Lie <- NaN
      
  for (i in 1:nrow(Data_All)){
    if(Data_All[i, "OrderOfLie"] == 1) {
          # Truth-First
          # Saturday == Truth
          # Sunday == Lie
            Data_All[i, "Person_Truth"] <- Data_All[i, "Sat_Person_NU"]
            Data_All[i, "Time_Truth"] <- Data_All[i, "Sat_Time_NU"]
            Data_All[i, "GPE_Truth"] <- Data_All[i, "Sat_GPE_NU"]
            Data_All[i, "Date_Truth"] <- Data_All[i, "Sat_Date_NU"]
            Data_All[i, "Total_Truth"] <- Data_All[i, "Total_Sat"]
      
            Data_All[i, "Person_Lie"] <- Data_All[i, "Sun_Person_NU"]
            Data_All[i, "Time_Lie"] <- Data_All[i, "Sun_Time_NU"]
            Data_All[i, "GPE_Lie"] <- Data_All[i, "Sun_GPE_NU"]
            Data_All[i, "Date_Lie"] <- Data_All[i, "Sun_Date_NU"]
            Data_All[i, "Total_Lie"] <- Data_All[i, "Total_Sun"]
    } else {
          # Lie-First
          # Saturday == Lie
          # Sunday == Truth
      
            Data_All[i, "Person_Truth"] <- Data_All[i, "Sun_Person_NU"]
            Data_All[i, "Time_Truth"] <- Data_All[i, "Sun_Time_NU"]
            Data_All[i, "GPE_Truth"] <- Data_All[i, "Sun_GPE_NU"]
            Data_All[i, "Date_Truth"] <- Data_All[i, "Sun_Date_NU"]
            Data_All[i, "Total_Truth"] <- Data_All[i, "Total_Sun"]
      
            Data_All[i, "Person_Lie"] <- Data_All[i, "Sat_Person_NU"]
            Data_All[i, "Time_Lie"] <- Data_All[i, "Sat_Time_NU"]
            Data_All[i, "GPE_Lie"] <- Data_All[i, "Sat_GPE_NU"]
            Data_All[i, "Date_Lie"] <- Data_All[i, "Sat_Date_NU"]
            Data_All[i, "Total_Lie"] <- Data_All[i, "Total_Sat"]
      
      } # END IF
    } # END LOOP

```

# Meaning of Column Headers
  1. `Participant.Nr.`    > unique participant number, corresponds with filenames of voice
                          tracks.
  2. `Condition.Nr.`      > Did participants talk about future or past weekend activities?
  3. `Location`           > Where the participant Data_All was collected
  4. `Date.Time`          > When the Data_All was collected
  5. `Age`                > Age of the participants during Data_Allcollection
  6. `Gender`             > Self-identified gender
  7. `Student..Y.N`       > Are the participants students?
  8. `English.Proficiency`  > On a level from 1 (not at all) to 5 (excellent) how do you 
                            rate your own level of English proficiency?
  9. `Ground.Truth.1`       > 1 Sentence Description of Past/Future Saturday activities
  10. `Ground.Truth.2`      > 1 Sentence Description of Past/Future Sunday activities
  11. `OrderOfLie`          > Did participants start by lying or telling the truth? 
                            Included as covariate, to detect any effect of modelling after 
                            the first description
  12. `Veracity.Accuracy`   > Was the 3rd researcher succesful in destinguishing the truth
  from the lie?
  13. `Veracity.Coding`     > Numerical values of 12
  14. `Judgement.By`        > Who provided the veracity judgement?
  15. `Frequency.of.Lie`    > Self-rated; on a level from 1 (never) to 10 (all the time), 
  how often do the activity that you lied about (e.g., how often do you go to the gym?).
  16. `Confidence.of.Deception` > On a scale from 1(not at all) to 10 (definitely) how sure
  are you that you have deceived the judge? In other words, higher scores indicate that the 
  participants believe their statements to be highly similar
  17. `Saturday`            > Transcriptions of the statement provided about the saturday
  18. `Sunday`              > Transcriptions of the statement provided about the sunday
  19. `NOTES`               > Any notable events that occurred during Data_All collection or 
  basis for the veracity judgements.
  20. `Transcriptions`      > Any notes regarding the transcriptions
  21. Named Entities (Sat_  = Saturday & Sun_ = Sunday)
      a. `_Veracity`        > Whether the description is a truth or a lie
      b. `_Person_NU`       > Non-unique person entities (people)
      c. `_Location_NU`     > Non-unique location entities (where)
      d. `_Time_NU`         > Non-unique time entities (smaller than a day)
      e. `_GPE_NU`          > Geographic locations
      f. `_Product_NU`      > Non-unique product entities (e.g., Nintendo)
      g. `_Date_NU`         > Non-unique date entities (e.g., monday)
      h. `_Money_NU`        > Non-unique money entities (e.g., 500 dollars)
  22. `Total_`              > Total number of named entities in truthful or deceptive statements
  23. `Predictive_Accuracy` > Veracity judgement made based on criteria that lies contain more details than truthful statements
      


# Predictive Accuracy

Confirmatory analysis indicated that lies, on average, contain more Named Entities than truthful statements. If we adopt this criteria as part of exploratory predictive modelling we find the following frequencies.

```{r Predictive Accuracy Before Exclusion, echo = FALSE}
# Create new column to store output
  Data_All$Predictive_Accuracy <- NaN

# Loop over all participants
  for (i in 1:nrow(Data_All)) {
    if (Data_All[i, "Total_Truth"] > Data_All[i, "Total_Lie"]){
      # truth longer than lie so decision rule not accurate
      Data_All[i, "Predictive_Accuracy"] <- "Mistaken"
    } else if (Data_All[i, "Total_Truth"] < Data_All[i, "Total_Lie"]) {
      # lie longer than truth so decision rule accurate
      Data_All[i, "Predictive_Accuracy"] <- "Correct"
    } else {
      # Lie and Truth of equal length so no decision can be made
      Data_All[i, "Predictive_Accuracy"] <- "Inconclusive"
    } # END IF 
  } # END LOOP

# Compute Frequencies
  accuracy <- table(Data_All$Predictive_Accuracy, Data_All$Time)
  accuracy <- cbind(accuracy, rowSums(accuracy))
  accuracy <- rbind(accuracy, colSums(accuracy))
  colnames(accuracy) <- c("Future", "Past", "Total")
  rownames(accuracy) <- c("Correct", "Inconclusive", "Mistaken", "Total")

print(accuracy)
```

The percentage accuracy attained from conclusive judgements (N = 50) is: 

```{R Percentage Accurate, echo = FALSE}
amount_deciscions <- (accuracy["Correct", "Total"] + 
                                   accuracy["Mistaken", "Total"])

accuracy_percentage <- round((accuracy["Correct", "Total"] / 
                                  amount_deciscions)*100,2) 

paste(accuracy_percentage, "%")
```

# Exclusion Criteria

Participants were excluded based on:

+ Inadherence to the instructions (N = 2)
    + PP. 5
    + PP. 21
+ Change of protocol (N = 15)
    + PP. 63 - PP. 77

```{R Exclusion Data}
Data_Excluded <- Data_All[-(c(5,21,63:77)),]
```

# Predictive Accuracy After Exclusion

```{R, echo = FALSE}
# Clean accuracy column 
Data_Excluded$Predictive_Accuracy <- NaN

# Loop over all participants
for (i in 1:nrow(Data_Excluded)) {
  if (Data_Excluded[i, "Total_Truth"] > Data_Excluded[i, "Total_Lie"]){
    # truth longer than lie so decision rule not accurate
    Data_Excluded[i, "Predictive_Accuracy"] <- "Mistaken"
  } else if (Data_Excluded[i, "Total_Truth"] < Data_Excluded[i, "Total_Lie"]) {
    # lie longer than truth so decision rule accurate
    Data_Excluded[i, "Predictive_Accuracy"] <- "Correct"
  } else {
    Data_Excluded[i, "Predictive_Accuracy"] <- "Inconclusive"
  } # END IF
} # END LOOP

# Compute Frequencies
  accuracy_excluded <- table(Data_Excluded$Predictive_Accuracy, Data_Excluded$Time)
  accuracy_excluded <- cbind(accuracy_excluded, rowSums(accuracy_excluded))
  accuracy_excluded <- rbind(accuracy_excluded, colSums(accuracy_excluded))
  colnames(accuracy_excluded) <- c("Future", "Past", "Total")
  rownames(accuracy_excluded) <- c("Correct", "Inconclusive", "Mistaken", "Total")

print(accuracy_excluded)

```
```{R Percentage Accurate After Exclusion, echo = FALSE}
amount_deciscions <- (accuracy_excluded["Correct", "Total"] + 
                                   accuracy_excluded["Mistaken", "Total"])

accuracy_excluded_percentage <- round((accuracy_excluded["Correct", "Total"] / 
                                  amount_deciscions)*100,0) 

paste(accuracy_excluded_percentage, "%")
```

# Visualization of Data
```{R echo=FALSE, message=FALSE, warning=FALSE}
# load package for creation of violin plots
library("yarrr")      
```

## Truthful Statements

```{R ViolingPlot Truths, echo = FALSE}
par(mfrow = c(1,2))

pirateplot(Data_All$Total_Truth ~ Data_All$Time, Data_All, pal = c("grey", "black"),
             xlab = "Time", ylab = "Amount of Named Entities Recognized",
             main = "Pre-Exclusion",
             family = "Times New Roman", gl.lwd = 0, inf.method = "se")

pirateplot(Data_Excluded$Total_Truth ~ Data_Excluded$Time, Data_Excluded, pal = c("grey", "black"),
             xlab = "Time", ylab = "Amount of Named Entities Recognized",
             main = "Post-Exclusion",
             family = "Times New Roman", gl.lwd = 0, inf.method = "se")

```

The amount of recognized named entities for truthful statements about past and future weekend activities. On the left the data is presented **before** exclusion, and on the right **after** exclusion of data. The black line represents the group mean, the box the standard error of the mean, and the spread of the data is outlined (Kampstra, 2008)

## Deceptive Statements

```{R ViolinPlot Lies, echo = FALSE}
par(mfrow = c(1,2))

pirateplot(Data_All$Total_Lie ~ Data_All$Time, Data_All, pal = c("grey", "black"),
             xlab = "Time", ylab = "Amount of Named Entities Recognized", 
              main = "Pre-Exclusion",
              family = "Times New Roman", gl.lwd = 0, inf.method = "se")

pirateplot(Data_Excluded$Total_Lie ~ Data_Excluded$Time, Data_Excluded, pal = c("grey", "black"),
             xlab = "Time", ylab = "Amount of Named Entities Recognized",
             main = "Post-Exclusion",
             family = "Times New Roman", gl.lwd = 0, inf.method = "se")


```
The amount of recognized named entities for deceptive statements about past and future weekend activities. On the left the data is presented **before** exclusion, and on the right **after** exclusion of data. The black line represents the group mean, the box the standard error of the mean, and the spread of the data is outlined (Kampstra, 2008)

```{R echo = FALSE}
# SAVE VIOLIN PLOTS

# png("violinplots_deceptive.png")
#   pirateplot(Data_Excluded$Total_Lie ~ Data_Excluded$Time, Data_Excluded, pal = c("grey", "black"),
#              xlab = "Time", ylab = "Amount of Named Entities Recognized",
#              main = "Deceptive Statements of Future and Past Activities",
#              family = "Times New Roman", gl.lwd = 0, inf.method = "se")
# 
# dev.off()
# 
# png("violingplots_truthful.png")
#  pirateplot(Data_Excluded$Total_Truth ~ Data_Excluded$Time, Data_Excluded, pal = c("grey", "black"),
#              xlab = "Time", ylab = "Amount of Named Entities Recognized",
#              main = "Truthful Statements of Future and Past Activities",
#              family = "Times New Roman", gl.lwd = 0, inf.method = "se")
# dev.off()
```

# Processed Data

The datafiles including newly created columns (e.g., total amount of NEs in truthful descriptions) are available as:

+ All collected data - [LieDetection_CombinedData_All.csv](https://github.com/SHogenboom/OnlineAppendix--LieDetection2017/blob/master/LieDetection_CombinedData_All.csv)
+ Data after excluding of participants - [LieDetection_Combined_Excluded.csv](https://github.com/SHogenboom/OnlineAppendix--LieDetection2017/blob/master/LieDetection_Combined_Excluded.csv)

```{R Save Data, echo = FALSE}
write.csv(Data_All, "LieDetection_CombinedData_All.csv")
write.csv(Data_Excluded, "LieDetection_Combined_Excluded.csv")
```

# Network Analysis

## Truthful Statements

```{R echo = FALSE}
require("psych")
require("qgraph")

nNames <- c("Person", "Time", "GPE", "Date")


pcor_truth_past <- partial.r(Data_Excluded[which(Data_Excluded[,"Time"] == "Past"), 
                                           c("Person_Truth", "Time_Truth", "GPE_Truth", "Date_Truth")])

pcor_truth_future <- partial.r(Data_Excluded[which(Data_Excluded[,"Time"] == "Future"), 
                                           c("Person_Truth", "Time_Truth", "GPE_Truth", "Date_Truth")])


par(mfrow = c(1,2))

truthful_network_past <- qgraph(pcor_truth_past, theme = "gray", layout = "circle", negDashed = TRUE,
       edge.labels = TRUE, edge.label.cex = 2, shape = "square", legend = FALSE, labels = nNames)

truthful_network_future <- qgraph(pcor_truth_future, theme = "gray", layout = truthful_network_past[["layout"]], 
                                  negDashed = TRUE, labels = nNames,
                                  edge.labels = TRUE, edge.label.cex = 2, shape = "square")

```

## Deceptive Statements

```{R echo=FALSE}

pcor_lie_past <- partial.r(Data_Excluded[which(Data_Excluded[,"Time"] == "Past")
                                         , c("Person_Lie", "Time_Lie", "GPE_Lie", "Date_Lie")])

pcor_lie_future <- partial.r(Data_Excluded[which(Data_Excluded[,"Time"] == "Future")
                                         , c("Person_Lie", "Time_Lie", "GPE_Lie", "Date_Lie")])

par(mfrow = c(1,2))

deceptive_network_past <- qgraph(pcor_lie_past, theme = "gray", layout = truthful_network_past[["layout"]], 
                            labels = nNames, negDashed = TRUE, legend = FALSE,
                            edge.labels = TRUE, edge.label.cex = 2, shape = "square")

deceptive_network_future <- qgraph(pcor_lie_future, theme = "gray", layout = truthful_network_past[["layout"]], 
                            labels = nNames, negDashed = TRUE,
                            edge.labels = TRUE, edge.label.cex = 2, shape = "square")

```


```{R CreateNetworkOutput, include = FALSE}
#  png("network_truths.png")
#    par(mfrow = c(1,2))
# 
#  truthful_network_past <- qgraph(pcor_truth_past, theme = "gray", layout = "circle", labels = nNames, negDashed = TRUE,
#         edge.labels = TRUE, edge.label.cex = 2, shape = "square", legend = FALSE)
# 
#  truthful_network_future <- qgraph(pcor_truth_future, theme = "gray", layout = truthful_network_past[["layout"]], labels = nNames, negDashed = TRUE,
#         edge.labels = TRUE, edge.label.cex = 2, shape = "square")
#  dev.off()
# 
# png("network_lies.png")
#    par(mfrow = c(1,2))
# 
#  deceptive_network_past <- qgraph(pcor_lie_past, theme = "gray", layout = truthful_network_past[["layout"]],
#                             labels = nNames, negDashed = TRUE, legend = FALSE,
#                              edge.labels = TRUE, edge.label.cex = 2, shape = "square")
# 
#  deceptive_network_future <- qgraph(pcor_lie_future, theme = "gray", layout = truthful_network_past[["layout"]],
#                              labels = nNames, negDashed = TRUE,
#                              edge.labels = TRUE, edge.label.cex = 2, shape = "square")
# dev.off()
```

